---
title: "The State of Out-of-district Political Donations in Canada"
author:
  - name: "Benedict Cummins-Mburu"
  - name: "Rohan Alexander"
    email: "rohan.alexander@utoronto.ca"
    attributes: 
        corresponding: true
abstract: "Political representation is fundamentally geographical, that is, in Canada Federal MPs represent geographical ridings. And yet, with various caveats, any Canadian living anywhere, can donate to any MP representing any riding. We examine the pattern of geographical political donations in Canada between 2013 and 2023. We document considerable out-of-riding donation patterns, especially differenced by an MPs race and gender. Our work has implications for thinking about the extent of political representation in Canada, and which parts of the population may be over- or under-represented."
thanks: "Code and data are available at: [https://github.com/bennypk1/geolocating-political-donations](https://github.com/bennypk1/geolocating-political-donations)."
date: today
date-format: long
format: pdf
linestretch: 1.3
number-sections: true
bibliography: references.bib
link-citations: true
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(knitr)
library(here)
library(grid)
```

# Introduction

# Literature Review

## Political Donation, Participation, and Representation

In Canada (and many other Western democratic nations) federal representation is fundamentally geographical: a Member of Parliament (MP) supposedly represents the aggregate views of their constituents, that is, everyone who lives within the boundaries of their riding. By voting during a general election, the citizens of a given electoral district increase the likelihood that their incumbent MP votes, acts, looks, and thinks how they want them to. Provided that (1) the division of people into ridings efficiently clusters like-minded Canadians, and (2) each Canadian participates equally (for example, through their one vote), we should expect a high degree of congruence between what a legislator does, and what their constituents want, achieving a truly representative democracy [@bednar_political_2011].

Beyond voting, individual citizens often seek to increase their local representation by volunteering at a local electoral district association (EDA), advocating for their preferred party online, or financially supporting a candidate’s campaign [@bednar_political_2011; @baker_getting_2016; @garnett_where_2024].Since these necessitate (1) more resources (e.g. time, effort, or money), (2) indicate greater political interest, and (3) often have larger impacts on electoral outcomes relative to voting, political scientists classically argue that these constitute “higher forms” of political participation [@bednar_political_2011].

Among these, donations stand out as a useful unit for (1) understanding the motivations behind political participation and (2) assessing whether higher forms of political participation distort the dyadic relationship between constituents and their MP. Here, we discuss two reasons why. First, donations are spatiotemporally unconstrained, telling us exactly when and where citizens seek representation [@gimpel_check_2008; @garnett_where_2024]. While the voter can only seek representation within their district and during an election cycle, the donor can legally contribute to various political entities, whenever, and as often as they want, subject to annual and per-election limits [@gimpel_check_2008; @canada_annual_2026]. Second, donations are the primary means by which individuals extend their representation beyond district borders (termed “surrogate” representation by [@gimpel_check_2008]). Enabled by modern digital communication, individuals are increasingly motivated to finance national parties, or non-local candidates, in an attempt to be represented on ideological, partisan, or identity-based grounds [@gimpel_check_2008]. For example, a Green-party partisan may realize their local candidate will lose, and are perfectly content funding another Green candidate in a closer race somewhere else. Similarly, a Black candidate may appeal to Black donors across the country seeking identity-based representation. Given that individual donors are the presently largest source of campaign funding [@carmichael_political_2014; @garnett_where_2024], high rates of out-of-district donation are of particular concern, since they risk shifting the effective representational structure away from one-Legislator-one-district and towards one-Legislator-one-donor base.

## Political Donations in the Canadian Context: What has been done

Much research has been done to determine the socio-demographic characteristics of donors and candidates in Canada. Most notably, Garnett analyzed candidate-bound donations during the 2015, 2019, and 2021 general elections, finding that donors are much more likely to originate from older, wealthier, and better-educated districts [@carmichael_political_2014; @garnett_where_2024]. The gender gap in giving has recieved particular attention, likely given the broader context of the established gender gap in political participation [see @tolley_who_2022 and @mcmahon_covid_2023]. Sevi compiled a dataset of demographic variables for all candidates who ran between 1867 to 2019, revealing, for example, that the share of women incumbents has dramatically increased since 1970 [@sevi_who_2021]. From 2004 to 2015, overall federal donation amounts have increased, particularly in Western Canada, and among Conservative political entities [@mcmahon_follow_2015]. The COVID-19 pandemic saw overall reduced political participation during it's first year [@mcmahon_covid_2023].

Garnett et al. propose that donors in Canada donate for two reasons: political (i.e. for representational gain) and transactional [@garnett_lifeblood_2022]. They find that political motivations are more prevalent overall, but especially among older and affluent donors who contributed larger sums. Transactional donations were more likely at the local level, while political donations were more likely at the national level [@garnett_lifeblood_2022].

When analyzing donor-recipient relationships, Canadian research has found that donors tend to donate more to candidates that they share aspects of their identity with [@tolley_who_2022; @besco_ethnic_2022]. Tolley et al. found that women donors are more likely to donate to women candidates [@tolley_who_2022], and Besco & Tolley found high rates of co-ethnic donation, particularly among South Asian donors [@besco_ethnic_2022]. More generally, South Asian political participation appears to be particularly heightened in Canada, relative to other ethnocultural groups [@garnett_where_2024]. Together with previous findings on donor motivation, these results the hypothesis that out-of-district donors in Canada are likely socially motivated, seeking individual-based representation.

Some work has also been done assessing the impacts of large amounts of political donation on representation. Among incumbent Legislators, voting behaviour is seemingly independent of the composition of their donor base [@peoples_impact_2008]. However, an older study found that local campaign donations during the 1993 and 1997 elections meaningfully affected the amount of votes a candidate receives [@eagles_effectiveness_2004].

Finally, we describe some unique aspects of Canadian political engagement, and particularly those that distinguish us from the U.S., where the large majority of work on out-of-district donations has been done. These differences will help us formulate hypotheses to guide our analysis. Most notably, party loyalty in Canada is weak ; it is estimated that only around 2% of Canadians are registered members of any political party [@garnett_lifeblood_2022], compared with around 40% in the U.S [@usafacts_how_2026]. This has been attributed to many things, including the argument that Canada's multi-party system de-emphasizes partisan membership [@carty_politics_2002]. This system also invites more ephemeral loyalties, based on specific policies or identities, and not broad-scope ideological affiliation. Many (historically) large parties in Canada were founded for specific reasons, for example the Bloc Quebecois, Reform Party, and Green Party, which cater(ed) to a very specific audience. High rates of immigration may have increase emphasis on cutural and not political belonging [@carty_politics_2002; @besco_ethnic_2022].

## American Out-of-District Donations to U.S. Congrssional Representatives

We were unable to find a rigorous treatment of out-of-district donations during Canadian federal elections. By contrast, American scholars have explicitly addressed out-of-district donations and their impacts on legislative representation, catalyzed primarily by Gimpel et al.’s extensive treatment of individual out-of-district donations during 4 Congressional elections from 1994 – 2004 [@gimpel_check_2008]. They propose the existence of “donor districts” ; a small set of older, wealthy, well-educated, and urban districts wherein a large majority of out-of-district donations originate, destined to Republican and Democrat candidates alike [@gimpel_check_2008; @bramlett_political_2011]. We should notice that these socio-economic factors are the same drivers of donor output in general across North America [@gimpel_political_2006; @garnett_where_2024]. Over this time period, they also observed an increase in the proportion of non-local donations received by the average candidate, from around 50% in 1994 to nearly 70% in 2004 [@gimpel_check_2008]. More recent studies suggest that this and similar trends have continued past 2004 [@bramlett_political_2011; @baker_getting_2016; @canes_wrone_out_2022]. Driven by the observation that the “closeness” of local races strongly predicted which districts received these non-local donations, Gimpel et al. argue that gaining a partisan advantage (and not seeking ideological or identity-based representation) is the primary motivation behind non-local donations [@gimpel_check_2008].

Follow-up American research has found that Out-of-district campaign contributions affects the voting behaviour of Congressional representatives, thereby distorting the supposedly-dyadic relationship between a representative and their district. A study of the 2006-2010 Congressional elections revealed that the average House representative becomes less responsive to their constituencies’ ideological preferences with increasing out-of-district public funding [@baker_getting_2016]. Canes-Wrone & Miller further explore the 2006-2016 Congressional elections to show that House voting behaviour is indeed associated with the preferences of their nation donor base [@canes_wrone_out_2022], contrasted against research finding that political action committees (U.S. equivalents to EDAs) do not affect roll call behaviour [@ansolabehere_why_2003]. Another study on Senatorial elections emphasizes the rivalrous relationship between constituents and out-of-district donors, experimentally showing that constituents who learned about outside donation expected their Senators to be less loyal to local interests [@nathan_donor_2025]. More generally, longitudinal analyses of out-of-state (and therefore out-of-district donations) during Senate elections echo many House findings [@sievert_out_2023; @jacobs_nationalization_2024; @keena_out_2024].

To an even greater extent than in Canada, individual donors are the primary source of funding for both U.S. campaigns and U.S. congressmen [@gimpel_check_2008; @baker_getting_2016]. This fact, paired with the many known effects of out-of-district donors on campaigns and representatives discussed above paint the picture that representatives “represent” little more than the whims of their national donor base.

# Data

## Data Overview

The data used in this paper was sourced primarily from Election Canada's Open Data website [@canada_open_2025]. Their (NAME OF SET) represents all donations made by individuals to registered federal political entities since January 2004 [@canada_open_2025]. Since that start date, the *Canada Elections Act* has required that all contributions (monetary or otherwise) valued at over \$20 be recorded by the recipient and shared with (WHO?) [@canada_understanding_2026]. Contributors associated with contributions valued at over \$200 \[representing the large majority of contributions ; CITATION\] must have their full name and address recorded as well [@canada_understanding_2026].

For donations valued at over \$200, the donor's name, location (including postal code), as well as the recipient's name, FED (if applicable), party (if applicable), type (e.g. National party, Candiate, etc.). To assess whether or not a donation was out-of-district, the publically available donor postal codes were matched with their corresponding FED using Statistics Canada's Postal Code Conversion Files. To simplify our cleaning and analysis, we restricted our focus to only contributions made within the 2013 representational order, corresponding to the period between Oct 2nd, 2015 and Apr 22nd, 2024 [@canada_redistribution_2025]. This is the largest such order present in the full dataset, notably capturing X% of total recorded contributions, and 3 general election periods [42nd, 43rd, and 44th respectively ; the same ones as @garnett_where_2024, allowing for direct comparison with her work]. Our data was further augnmented with CENSUS DATA ... Our cleaning procedure is described in full in Appendix X.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-1-1
#| tbl-cap: "First 4 Rows of Cleaned Individual Donations Data (Subset of Columns)."
Table_1_1 <- read_csv(here::here("other/tables/Table_1_1.csv"))
Table_1_1 <- Table_1_1 %>% head(4)
Table_1_1 %>% knitr::kable()
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-1-2
#| tbl-cap: "First 4 Rows of Cleaned District Flows Data (Subset of Columns)"
Table_1_2 <- read_csv(here::here("other/tables/Table_1_2.csv"))
Table_1_2 <- Table_1_2 %>% head(4)
Table_1_2 %>% knitr::kable()
```

Our cleaning produced two related datasets that were used for different parts of the analysis. The first represented all individual contributions made during our focal period, to any recipient. @tbl-1-1 displays the first few rows and columns of this data. This dataset was used to document prelimiary findings, and understand how out-of-district donations relate to in-district donations. The second is a subset of the first, and includes only donations whose recipients are local political entities (Candidates, Nomination Contestants, and EDAs). This second dataset was subsequently wrangled such that each row represented a donor-recipient district pair, allowing us to effecrively track interdistrict flows of donations at various points in time and space. @tbl-1-2 displays the first few rows and columns of this data. National entities (Leadership Contestants and Registered Parties) were necessarily excluded from this data since they cannot be tied to any one local FED.

## Individual Donations Data Summary

Below we present a few longitudinal visualizations of the first dataset. We start with the disclaimer that all types of donations occurred at much higher rates in December compared to other months, regardless of year, recipient type, party affiliation, etc. This is likely a clerical error (ELABORATE). This did not impact our analysis (CHECK THIS WHEN FLOW ANALYSIS IS CARRIED OUT), however December donations in all Figures presented below were imputed using their neighboring November and January values, and their recorded monetary amounts were redistributed accross their associated years so that nothing was lost. This served to reduce visual clutter from large, phantom peaks at the end of each year.

Although Figures below only display trends in Monthly donation sums, donation counts were largely identical, and are shown in Appendix X.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-national-entity
#| fig-cap: "Federal Political Donations to National Entities (Oct 2015 - Apr 2024), Subsetted by Entity. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively."
Figure_1 <- readRDS(here::here("other/figures/Figure_1a.rds"))
Figure_1 + 
  labs(
    title = NULL,
    fill = NULL
    ) +
  theme(
    axis.title.y = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.title = NULL,
    legend.key.size = grid::unit(5, "mm")
  )
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-national-party
#| fig-cap: "Federal Political Donations to National Entities (Oct 2015 - Apr 2024), Subsetted by Party. Only data from top 4 parties is shown. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively."
Figure_4 <- readRDS(here::here("other/figures/Figure_4a.rds"))
Figure_4 + 
  labs(
    title = NULL,
    fill = NULL
    ) +
  theme(
    axis.title.y = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.title = NULL,
    legend.key.size = grid::unit(5, "mm")
  )
```

@fig-national-entity and @fig-national-party display the timeline of donatios to Leadership Contestants and National (i.e. Registered) parties. The 6 main peaks in the graph clearly correspond to general election periods (peaks followed immediately by grey lines) and Conservative party leadership campaigns (blue-tipped peaks in @fig-national-entity). Interestingly, elections and leadership contests result in relatively similar amounts of donations to national political entites, with the notable exception of the 2022 Conservative leadership race, won by current Opposition Leader Pierre Poilievre. Overall, donations to national entities seem to have increased during this period, following the increasing trend observed from 2004 - 2015 by McMahon and Wolfe [@mcmahon_follow_2015], and particularly starying in 2022, a recent phenomenon driven primarily by increased donations to the Conservative and Liberal parties (@fig-national-party). Although National Parties are somewhat out of scope when analyzing out-of-district donations, understanding their trends provides helpful context, especially since they represent the large majority (\~90%) of donations in our dataset.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-local-in-entity
#| fig-cap: "In-District Donations to Local Political Entities (Oct 2015 - Apr 2024), Subsetted by Entity. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively."
Figure_2 <- readRDS(here::here("other/figures/Figure_2a.rds"))
Figure_2 + 
  labs(
    title = NULL,
    fill = NULL
    ) +
  theme(
    axis.title.y = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.title = NULL,
    legend.key.size = grid::unit(5, "mm")
  )
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-local-out-entity
#| fig-cap: "Out-of-District Donations to Local Political Entities (Oct 2015 - Apr 2024), Subsetted by Entity. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively."
Figure_3 <- readRDS(here::here("other/figures/Figure_3a.rds"))
Figure_3 + 
  labs(
    title = NULL,
    fill = NULL
    ) +
  theme(
    axis.title.y = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.title = NULL,
    legend.key.size = grid::unit(5, "mm")
  )
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-local-out-party
#| fig-cap: "Out-of-District Donations to Local Political Entities (Oct 2015 - Apr 2024), Subsetted by Party. Only data from top 4 parties is shown. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively."
Figure_5 <- readRDS(here::here("other/figures/Figure_5a.rds"))
Figure_5 + 
  labs(
    title = NULL,
    fill = NULL
    ) +
  theme(
    axis.title.y = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.title = NULL,
    legend.key.size = grid::unit(5, "mm")
  )
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-local-in-party
#| fig-cap: "In-District Donations to Local Political Entities (Oct 2015 - Apr 2024), Subsetted by Party. Only data from top 4 parties is shown. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively."
Figure_6 <- readRDS(here::here("other/figures/Figure_5a.rds"))
Figure_6 + 
  labs(
    title = NULL,
    fill = NULL
    ) +
  theme(
    axis.title.y = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.title = NULL,
    legend.key.size = grid::unit(5, "mm")
  )
```

@fig-local-in-entity, @fig-local-out-entity, @fig-local-in-party, and @fig-local-out-party display various subsets of donations to Local Political Entities (i.e. EDAs, Candiates, and Registered Associations). @fig-local-in-entity and @fig-local-in-party showcase in-district local donations, broken down by entity and party respectively. @fig-local-out-entity and @fig-local-out-party show the same breakdowns but for out-of-district local donations (the reader should note that donations to National political bodies cannot be "in-district" or "out-of-district", hence why it is not considered at that level). Like with Figures @fig-national-entity and @fig-national-party, donations to EDAs and Candidates peak during general election periods. Nomination Contestants are relatively scarce, even though many nomination capmaigns occured in Canada during this period (CITATION). This is likely because these contests recieve little media coverage, and occur entirely among local by the party [@pruysers_nomination_contests_2016]. This is unlike Leadership contests, which receive widespread national media coverage, at least among the largest parties (CITATION), whichm may explain why they appear clearly in @fig-national-entity. Donations to Candidates are restricted to elections (HOWEVER THIS MAY BE AN ISSUE WITH THE DATA, TO BE INVESTIGATED MORE THOROUGHLY). Most notably, the longitudinal distribution of in and out-of-district donations is seemingly identical, both during and outside of general elections.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: table-prop-ood-entity-out
#| tbl-cap: "Summary Statistics for Donations given outside election cycles."
Table_2_1 <- read_csv(here::here("other/tables/Table_2_1.csv"))
Table_2_1 <- Table_2_1 %>% filter(str_detect(political_entity, "^Registered")) # TO FOCUS ON MAIN POINT
Table_2_1 <- Table_2_1 %>%
  mutate(across(everything(), ~replace(.x, .x == 0, NA)))
Table_2_1 %>% knitr::kable()
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: table-prop-ood-entity-in
#| tbl-cap: "Summary Statistics for Donations given within election cycles."
Table_2_2 <- read_csv(here::here("other/tables/Table_2_2.csv"))
Table_2_2 <- Table_2_2 %>% head(3) # TO FOCUS ON MAIN POINT
Table_2_2 <- Table_2_2 %>%
  mutate(across(everything(), ~replace(.x, .x == 0, NA)))
Table_2_2 %>% knitr::kable()
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: table-in-out-cycle
#| tbl-cap: "Donations in and out of election cycles by party, for the top 4 Parties."
Table_2_3 <- read_csv(here::here("other/tables/Table_2_3.csv"))
Table_2_3 <- Table_2_3 %>%
  mutate(across(everything(), ~replace(.x, .x == 0, NA)))
Table_2_3 %>% knitr::kable()
```

@table-prop-ood-entity-out, and @table-prop-ood-entity-in, and @table-in-out-cycle showcase some key annual summary statistics. Accross parties, donations are mainly accumulated outside of election capmaigns (see @table-in-out-cycle) ; around 10-15% of donor-sourced funds among the top 4 Parties, with the notable exception of the NDP at 25%. However, roughly 45% of donor-sourced funds to local entites (i.e. EDAs and Candidates) orginated from outside their districts, both in and out of election cycles. The identical and stagnant longitudinal distribution between in and ou-of donors illustrated in Figures @fig-local-out-entity shows that this proportion is roughly constant over time, unlike the increasing OOD proportion of donations in the U.S. over a similar time period (CITATION). In all cases, the count proportions were smaller than the amount proportions. This means that, on average, out-of-district and out-of-cycle donors are contributing larger amounts.

## District Flows Data Summary

#TODO

\newpage

\appendix
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{table}{0}

# Background on Federal Donation Limits {#sec-background}

The Canadian Elections Act imposes contribution limits on the following registered, federal political entities [see a full list of definitions at @canada_political_2026]: political parties, local electoral district associations (EDAs), candidates, party nomination contestants, and party leadership contestants [@canada_annual_2026]. Regulated expenses for any of the entities themselves cannot exceed specific values (determined on an annual basis, different for each riding. Leadership contestants are a notable exception, with no campaign spending limits. See @canada_expenses_2026). Notably, registered third parties are not contribution-limited by the Act [@canada_third_2026]. However, expenses relating to their activities are capped on a per-riding basis, similar to other political entities [@canada_expenses_2026]

Only Canadian citizens or permanent residents can make contributions [@canada_understanding_2026]. Any donor who contributed \$20 or more to any of the five federal political entities listed above must have their name recorded (but not reported to Elections Canada ; @canada_understanding_2026). Donors with contributions exceeding \$200 must also have their names and addresses recorded and reported to Elections Canada [@canada_understanding_2026]. Cash donations above \$20 are prohibited [@canada_understanding_2026].

| Political Entity | 2026 Annual Limit (CAD) |
|------------------------------------------------------|------------------|
| To each registered national party | 1,775 |
| In total, to all EDAs, nomination contestants, and candidates (for each registered party) | 1,775 |
| In total, to each independent or non-affiliated candidate | 1,775 |

: 2026 Annual Limits on Individual Contributions, Loans, and Loan Guarantees (adapted from Elections Canada website).

| Politician            | Own-Campaign Contribution Limit |
|-----------------------|---------------------------------|
| Nomination Contestant | 2,775                           |
| Candidate             | 5,000                           |
| Leadership Contestant | 25,000                          |

: 2026 Special Cases for Politicians

**Other Info to put somewhere:**

- Electoral district boundaries are subject to change every 10 years, as required by the Constitution of Canada [@canada_redistribution_2025].

- Increases by \$25 each year

# Source to IJF {#sec-source}

The raw donations dataset is from Elections Canada's "Open data on contributions" page (see @canada_open_2025), and was accessed for this study on May 5, 2026. This dataset represents all contribution reports filed by registered political entities (see Appendix A) from January 2004 to present [@canada_open_2025]. This is the exact data the IJF starts with.

Table B1 shows how each source column is cleaned and mapped by the IJF. Source data had 10,095,198 entries, while IJF had 8,095,810 entries.

IJF appended extra archived data from the site [see @canada_open_2025], covering donations to political entities and candidates from 1993 - 2004. These do not matter for our analysis, and were discarded.

In the raw data, 31,873 rows (0.3% of the total data) is not from Individual donors (Buisnesses, Corporations, Trade, Associations, Government, Unincorporated). These can immediately be ignored.

# IJF to Clean {#sec-ijfclean}

# References