---
title: "Out-of-district Political Donations in Canada (2015 - 2024): A Descriptive Analysis"
author:
  - name: "Benedict Cummins-Mburu"
  - name: "Rohan Alexander"
    email: "rohan.alexander@utoronto.ca"
    attributes: 
        corresponding: true
thanks: "Code and data are available at: [https://github.com/bennypk1/geolocating-political-donations](https://github.com/bennypk1/geolocating-political-donations)."
abstract: |
  Federal political representation in Canada is fundamentally geographical; one Members of Parliament (MP) represents one electoral district. Despite this representational structure, Canadians can generally donate to a host of political entities, regardless of proximity. Using an open-source political financing dataset compiled by Election Canada, we analyze the extent, timing, and flow patterns of out-of-district donations in Canada, from 2015 to 2024. We find [...]. These and other findings help to [...] and more generally in a critical assessment of effective representation in a Westminster-style parliamentary democracy. \

  **Keywords:** Canadian politics; out-of-district donations; surrogate representation; political participation; political geography
date: today
date-format: long
format: pdf
linestretch: 1.3
number-sections: true
bibliography: references.bib
link-citations: true
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(tinytable)
library(knitr)
library(here)
library(grid)
```

# Introduction {#sec-intro}

# Literature Review {#sec-lit}

## Political Donation, Participation, and Representation {#sec-lit-1}

In Canada and many other Western democratic nations, federal representation is fundamentally geographical: a Member of Parliament (MP) is a representative of everyone who lives within the boundaries of their riding. By voting during a general election, the citizens of a given electoral district increase the likelihood that their incumbent MP votes, acts, looks, and thinks how they want them to. Provided that: (1) the division of people into ridings efficiently clusters like-minded Canadians, and (2) each Canadian participates equally (for example, through their one vote), we should expect a high degree of congruence between what a legislator does, and what their constituents want [@bednar_political_2011].

Beyond voting, individual citizens can increase their local representation by volunteering at a local electoral district association (EDA), advocating for their preferred party online, or financially supporting a candidate’s campaign [@bednar_political_2011; @baker_getting_2016; @garnett_where_2024]. Since these necessitate more resources (e.g. time, effort, or money), indicate greater political interest, and often have larger impacts on electoral outcomes relative to voting, these constitute “higher forms” of political participation [@bednar_political_2011].

Among these, donations stand out as a useful unit for understanding the motivations behind political participation and assessing whether higher forms of political participation distort the dyadic relationship between constituents and their MP. Here, we discuss two reasons why. First, donations are spatiotemporally unconstrained, telling us exactly when and where citizens seek representation [@gimpel_check_2008; @garnett_where_2024]. While a voter can only seek representation within their district and during an election cycle, a donor can legally contribute to various political entities, whenever, and as often as they want, subject to annual and per-election limits [@gimpel_check_2008; @canada_annual_2026]. Second, donations are the primary means by which individuals extend their representation beyond district borders. Enabled by modern digital communication, individuals are increasingly motivated to finance national parties, or non-local candidates, in an attempt to be represented on ideological, partisan, or identity-based grounds [@gimpel_check_2008]. For example, a Green party partisan may realize their local candidate will lose, and are content funding another Green candidate in a closer race somewhere else. Similarly, a Black candidate may appeal to Black donors across the country seeking identity-based representation. Given that individual donors are the largest source of campaign funding [@carmichael_political_2014; @garnett_where_2024], high rates of out-of-district donation are of particular concern, since they risk shifting the effective representational structure away from one-Legislator-one-district and towards one-Legislator-one-donor base.

## Political Donations in the Canadian Context: What has been done {#sec-lit-2}

Much research has been done to determine the socio-demographic characteristics of donors and candidates in Canada. Most notably, an analysis of candidate-bound donations during the 2015, 2019, and 2021 general elections found that donors are more likely to originate from older, wealthier, and better-educated districts [@carmichael_political_2014; @garnett_where_2024]. The gender gap in giving has recieved particular attention, likely given the broader context of the established gender gap in political participation [see @tolley_who_2022 and @mcmahon_covid_2023]. From 2004 to 2015, overall federal donation amounts have increased, particularly in Western Canada, and among Conservative political entities [@mcmahon_follow_2015]. The COVID-19 pandemic saw overall reduced political participation [@mcmahon_covid_2023].

@garnett_lifeblood_2022 propose that donors in Canada donate for two reasons: political (i.e. for representational gain) and transactional (i.e. a favour for a friend, or a willingness to be seen). They find that political motivations for giving are more prevalent overall, but especially among older and affluent donors who contributed larger sums. Transactional donations were more likely at the local level, while political donations were more likely at the national level.

Canadian donors tend to donate more to candidates that they share aspects of their identity with. For instance, women donors are more likely to donate to women candidates [@tolley_who_2022], and there are high rates of co-ethnic donation, particularly among South Asian donors [@besco_ethnic_2022]. More generally, South Asian political participation appears to be particularly heightened in Canada, relative to other ethnocultural groups [@garnett_where_2024]. Together with previous findings on donor motivation, these results suggest that out-of-district donors in Canada are predominantly socially motivated, seeking individual-based representation.

Some work has also been done assessing the impacts of large amounts of political donation on representation. Among incumbent legislators, voting behaviour is seemingly independent of the composition of their donor base [@peoples_impact_2008]. However, @eagles_effectiveness_2004 found that local campaign donations during the 1993 and 1997 elections meaningfully affected the number of votes a candidate receives.

Finally, we describe some unique aspects of Canadian political engagement, and particularly those that distinguish Canada from the U.S., where the large majority of work on out-of-district donations has been done. These differences will help us formulate hypotheses to guide our analysis. Most notably, party loyalty in Canada is weak; it is estimated that only around 2% of Canadians are registered members of any political party [@garnett_lifeblood_2022], compared with around 40% in the U.S. [@usafacts_how_2026]. This has been attributed to many things, including the argument that Canada's multi-party system de-emphasizes partisan membership [@carty_politics_2002]. This system also invites more ephemeral loyalties, based on specific policies or identities, and not broad-scope ideological affiliation. Many (historically) large parties in Canada were founded for specific reasons, for example the Bloc Québécois, Reform Party, and Green Party, which cater(ed) to a very specific audience. High rates of immigration may have increase emphasis on cutural and not political belonging [@carty_politics_2002; @besco_ethnic_2022].

## American Out-of-District Donations to U.S. House Representatives {#sec-lit-3}

In the U.S., @gimpel_check_2008 examine out-of-district donations and their impacts on legislative representation, looking at individual out-of-district donations during 4 Congressional elections from 1994 – 2004. They propose the existence of “donor districts”; a small set of older, wealthy, well-educated, and urban districts wherein a large majority of out-of-district donations originate, destined to Republican and Democrat candidates alike [@gimpel_check_2008; @bramlett_political_2011]. We should notice that these socio-economic factors are the same drivers of donor output in general across North America [@gimpel_political_2006; @garnett_where_2024]. Over this time period, they also observed an increase in the proportion of non-local donations received by the average candidate, from around 50% in 1994 to nearly 70% in 2004 [@gimpel_check_2008]. More recent studies suggest that this and similar trends have continued past 2004 [@bramlett_political_2011; @baker_getting_2016; @canes_wrone_out_2022]. Driven by the observation that the “closeness” of local races strongly predicted which districts received these non-local donations, Gimpel et al. argue that gaining a partisan advantage (and not seeking ideological or identity-based representation) is the primary motivation behind non-local donations [@gimpel_check_2008].

Follow-up American research has found that out-of-district campaign contributions affects the voting behaviour of Congressional representatives, thereby distorting the supposedly-dyadic relationship between a representative and their district. A study of the 2006-2010 Congressional elections revealed that the average House representative becomes less responsive to their constituencies’ ideological preferences with increased out-of-district public funding [@baker_getting_2016]. @canes_wrone_out_2022 further explore the 2006-2016 Congressional elections to show that House voting behaviour is indeed associated with the preferences of their nation donor base, contrasted against research finding that political action committees (U.S. equivalents to EDAs) do not affect roll call behaviour [@ansolabehere_why_2003]. Another study on senatorial elections emphasizes the rivalrous relationship between constituents and out-of-district donors, experimentally showing that constituents who learned about outside donation expected their senators to be less loyal to local interests [@nathan_donor_2025]. More generally, longitudinal analyses of out-of-state (and therefore out-of-district donations) during Senate elections echo many House findings [@sievert_out_2023; @jacobs_nationalization_2024; @keena_out_2024].

To an even greater extent than in Canada, individual donors are the primary source of funding for both U.S. campaigns and U.S. congressmen [@gimpel_check_2008; @baker_getting_2016]. This fact, paired with the many known effects of out-of-district donors on campaigns and representatives discussed above paint the picture that representatives “represent” little more than the whims of their national donor base.

# Data {#sec-data}

## Data Overview {#sec-data-1}

The data used in this paper was originally sourced from Election Canada's Open Data website [@canada_open_2025]. Since 2004, the *Canada Elections Act* has required that all contributions (monetary or otherwise) valued at over \$20 be recorded by the recipient and shared with the Government [@canada_understanding_2026]. The Elections Canada dataset on donations represents all donations made by individuals to registered federal political entities since that date [@canada_open_2025]. Additionally, contributors who contributed over \$200 must have their full name and address (including their postal code) recorded as well [@canada_understanding_2026]. For all contributions, the recipient's name, district (if applicable), party (if applicable), type (e.g. National party, Candiate, etc.), contribution amount (we used the sum of monetary and non-monetary contributions), and the date the contribution was received is also recorded. The data we used is a processed version of the Elections Canada dataset, made available to us by the Investigative Journalism Foundation (IJF). Their cleaning procedure is elaborated on further in **Appendix X**.

Every 10 years, federal electoral district (FED) boundaries are redrawn to account for population changes that occured within the decade [@canada_redistribution_2025]. We restricted our scope to contributions made during the 2013 representational order, effective between Oct 2nd, 2015 and Apr 22nd, 2024 [@canada_redistribution_2025]. This is the largest such order present in the full dataset, capturing **X > 60%** of all recorded contributions since 2004, and spanning 3 general election periods (42nd, 43rd, and 44th). This decision allowed us to focus on a single, static set of FEDs, greatly simplifying our cleaning procedures and overall analysis. To assess whether or not a donation was out-of-district, donor postal codes were matched with their corresponding FED using Statistics Canada's Postal Codes by Federal Riding File, for the 2013 representational order, made accessible by the University of Toronto [@statcan_pcfrf_2022]. As mentioned above, donor postal codes are only available for constributions larger than $200, however these represent a large proportion (**X > 90%**) of the toal contributed amount during the period we considered. Our dataset was further augmented with demographic statistics (i.e. average wage, dominant ethnicity, etc.) for each FED Using the 2021 Canadian Census. A detailed description of our data cleaning and augmentation process and decisions made can be found in **Appendix X**.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-1-1
#| tbl-cap: "Variables used in our study, with examples and descriptions."
read_csv(here::here("other/tables/Table_1_3.csv")) %>% 
  tinytable::tt(width = c(0.2, 0.2, 0.6)) %>%
  format_tt(escape = TRUE)
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-1-2
#| tbl-cap: "First 4 Rows of Cleaned District Flows Data (Subset of Columns)"
read_csv(here::here("other/tables/Table_1_2.csv")) %>% 
  tinytable::tt() %>% 
  format_tt(escape = TRUE)
```

@tbl-1-1 provides a description of the focal variables used throughout this study, and an example of each. Most contributions included in our analyses (**X~88%**) were addressed to national-level political entities; either party leadership contestants, or the parties themselves. Since these entities are not tied to local FEDs, contributions addressed to them cannot be declared as in or out of district. Longitudinal trends in these contributions are briefly discussed in @sec-data-2. To explicitly track large-scale flows between donors and local political entities (candidates, nomination contestants, and electoral district associations), we wrangled a secondary dataset representing aggregated flows between every pair of FEDs. Flows were recorded (e.g. grand total aomunt sent, total amount sent during an election cycle, etc.). The first few rows and columns of this dataset are shown in @tbl-1-2.

## General Trends in Federal Contributions {#sec-data-2}

We start by briefly documenting summaries and trends in donatons generally, to provide context for our analysis of out-of-district donations.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-2-1
#| fig-cap: "Distribution of contributions by recipient entity (Oct 2015 - Apr 2024), shown during (a) and outside (b) election cycles. Contributions to Nomination contestants were negligibly small, and removed to simplify the visualization."
readRDS(here::here("other/figures/Figure_2_1.rds"))
```

@fig-2-1 displays the breakdown of total amount contributed by type of recipient (i.e. political entity) and period (in or out of an election cycle). Donations recived during election cycles represent a minority of all donations (**X~20%**), and local and national-level entities receive similar amounts of donor-based funding. Local funding is shared evenly between candidates and registered (i.e. district) associations. Outside of election cycles, national entities (primarily parties but also a non-negligible sum to leadership contestants) recieve sustantially more donations than local ones. During electoral downtime, notable amounts of local funding are raised only by district associations ; donations to electoral candidates are effectively contrained to campaign periods. Both in and out of election cycles, nomination contestants received relatively tiny amounts of donations, and thus are not shown. Funding distributions are similar when counts are displayed instead of amounts, and when specific events are examined (e.g. 42nd general election, or the period between the 4nd and 43rd general elections. See **Appendix X** for such figures).

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-2-2-v1
#| fig-cap: "Time series of federal political donations from Oct 2015 - Apr 2024. Trends are seperated by the political party of the recipient. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively. Version 1."
readRDS(here::here("other/figures/Figure_2_2_v1.rds"))
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-2-2-v2
#| fig-cap: "Time series of federal political donations from Oct 2015 - Apr 2024. Trends are seperated by the political party of the recipient. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively. Version 2."
readRDS(here::here("other/figures/Figure_2_2_v2.rds"))
```

@fig-2-2-v1 (and @fig-2-2-v2) display the timeline of donations during our study period, separated by the party of the recipient. We briefly note that, for this visualization, donations dated in December of each year were redistributed accross the other months of that year to address a clerical error in the original dataset (see **Appendix X** for more details). The 6 main peaks in the graph clearly correspond to general election periods (peaks followed immediately by grey lines) and Conservative party leadership campaigns (peaks in the blue line in @fig-2-2-v2). Interestingly, elections and leadership contests result in relatively similar amounts of donations, with the notable exception of the 2022 Conservative leadership race, won by current Opposition Leader Pierre Poilievre. Overall, donations seem to have increased during the study period, following the increasing trend observed from 2004 - 2015 by McMahon and Wolfe [@mcmahon_follow_2015], and particularly starying in 2022, a recent phenomenon driven primarily by increased donations to the Conservative and Liberal parties.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-2-4-v1
#| fig-cap: "Time series of federal political donations to in-district local entities (Candidates or Electoral District Associations), from Oct 2015 - Apr 2024. Trends are seperated by the political party of the recipient. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively. Version 1."
readRDS(here::here("other/figures/Figure_2_4_v1.rds"))
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-2-5-v1
#| fig-cap: "Time series of federal political donations to out-of-district local entities (Candidates or Electoral District Associations), from Oct 2015 - Apr 2024. Trends are seperated by the political party of the recipient. Dashed vertical lines indicate general election dates for the 42nd, 43rd, and 44th general elections respectively. Version 1."
readRDS(here::here("other/figures/Figure_2_5_v1.rds"))
```

Since the trends in @fig-2-2-v1 are dominated by national donations, we also display the timeline of contributions to select local entities (Candidates and EDAs) in @fig-2-4-v1 and @fig-2-5-v1. @fig-2-4-v1 shows the timeline of federal donations to local entities from donors located within the same electoral district, while @fig-2-5-v1 plots the same metrics for out-of-district donors. As expected from @fig-2-2-v1, donations to EDAs and Candidates peak during general election periods. Most notably, the longitudinal distribution of in and out-of-district donations is seemingly identical, both during and outside of general elections, and accross the main political parties.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-2-1
#| tbl-cap: "Rates of funding sourced from out-of-district donors among local political entities."
read_csv(here::here("other/tables/Table_2_1.csv")) %>% 
  tinytable::tt(width = 1) %>% 
  format_tt(escape = TRUE)
```

@tbl-2-1 showcases some key summary statistics for local political entities. Both in and out of election cycles, candidates and EDAs receive roughly half of their donor-sourced funding from outside their district. Given how similar in-district and out-of-district contribution trends are over time (@fig-2-4-v1 and @fig-2-5-v1), the annual means presented in @tbl-2-1 are, at first glance, representative of the proportions at any specific period of time. While the out-of-district funding proportion in the U.S. has dramatically increased over time, we observe no evidence of this in the Canadian context. In all cases presented, the proportion of out-of-district contributions (OOD % (Count) in @tbl-2-1) is smaller than the proportion of funds recieved from out-of-district (OOD % (Sum) in @tbl-2-1). This tells us that non-local donors are, on average, contributing larger sums of money to these political entities.

Through this preliminary analysis, we discover that consistent national-level donations dominate trends in contributions over time ; party leadership contests create swells of contribution, comparable in size to those of general elections, and the top 3 parties claim both a majority of donors, and of total contributed amount. Despite this, local EDAs (and candidates during election periods) form a consistently non-negligible fraction of donation recipients, even outside of election periods. Further, nearly half of the funds raised by EDAs and/or candidates originate from donors located outside the boundaries of their respective districts, both during and outside of election cycles (@tbl-2-1). Together these findings help to situate further analysis of interdistrict flows within the broader context of federal donations as a whole.

## Interdistrict Flows {#sec-data-3}

Here we document general trends in interdistrict flows through space and time.


\newpage

\appendix
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{table}{0}

# Background on Federal Donation Limits {#sec-background}

The Canadian Elections Act imposes contribution limits on the following registered, federal political entities [see a full list of definitions at @canada_political_2026]: political parties, local electoral district associations (EDAs), candidates, party nomination contestants, and party leadership contestants [@canada_annual_2026]. Regulated expenses for any of the entities themselves cannot exceed specific values (determined on an annual basis, different for each riding. Leadership contestants are a notable exception, with no campaign spending limits. See @canada_expenses_2026). Notably, registered third parties are not contribution-limited by the Act [@canada_third_2026]. However, expenses relating to their activities are capped on a per-riding basis, similar to other political entities [@canada_expenses_2026]

Only Canadian citizens or permanent residents can make contributions [@canada_understanding_2026]. Any donor who contributed \$20 or more to any of the five federal political entities listed above must have their name recorded (but not reported to Elections Canada ; @canada_understanding_2026). Donors with contributions exceeding \$200 must also have their names and addresses recorded and reported to Elections Canada [@canada_understanding_2026]. Cash donations above \$20 are prohibited [@canada_understanding_2026].

| Political Entity | 2026 Annual Limit (CAD) |
|------------------------------------------------------|------------------|
| To each registered national party | 1,775 |
| In total, to all EDAs, nomination contestants, and candidates (for each registered party) | 1,775 |
| In total, to each independent or non-affiliated candidate | 1,775 |

: 2026 Annual Limits on Individual Contributions, Loans, and Loan Guarantees (adapted from Elections Canada website).

| Politician            | Own-Campaign Contribution Limit |
|-----------------------|---------------------------------|
| Nomination Contestant | 2,775                           |
| Candidate             | 5,000                           |
| Leadership Contestant | 25,000                          |

: 2026 Special Cases for Politicians

**Other Info to put somewhere:**

- Electoral district boundaries are subject to change every 10 years, as required by the Constitution of Canada [@canada_redistribution_2025].

- Increases by \$25 each year

# Source to IJF {#sec-source}

The raw donations dataset is from Elections Canada's "Open data on contributions" page (see @canada_open_2025), and was accessed for this study on May 5, 2026. This dataset represents all contribution reports filed by registered political entities (see Appendix A) from January 2004 to present [@canada_open_2025]. This is the exact data the IJF starts with.

Table B1 shows how each source column is cleaned and mapped by the IJF. Source data had 10,095,198 entries, while IJF had 8,095,810 entries.

IJF appended extra archived data from the site [see @canada_open_2025], covering donations to political entities and candidates from 1993 - 2004. These do not matter for our analysis, and were discarded.

In the raw data, 31,873 rows (0.3% of the total data) is not from Individual donors (Buisnesses, Corporations, Trade, Associations, Government, Unincorporated). These can immediately be ignored.

# IJF to Clean {#sec-ijfclean}

# Modifications to Data for Figures

- LONGITUDINAL FIGURES: We start with the disclaimer that all types of donations occurred at much higher rates in December compared to other months, regardless of year, recipient type, party affiliation, etc. This is likely a clerical error (ELABORATE). This did not impact our analysis (CHECK THIS WHEN FLOW ANALYSIS IS CARRIED OUT), however December donations in all Figures presented below were imputed using their neighboring November and January values, and their recorded monetary amounts were redistributed accross their associated years so that nothing was lost. This served to reduce visual clutter from large, phantom peaks at the end of each year.


# References